---
title: "Data cleaning"
author: "Jae Yeon Kim"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: true
    theme: united
---

# Setup 

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
        purrr, # for functional programming
        foreign, # for importing stata files 
        tidyverse, # for the tidyverse framework 
        psych, # for psych tools  
        ggthemes, # for fancy ggplot themes
        here, # for self-contained projects
        ggpubr, # for pub-ready themes 
        glue # pasting string and objects
)

source(here("functions", "utils.r"))

# for publication-friendly theme 
theme_set(theme_pubr())
```

# Import files 

The data has 406 variables (the number of columns) and 6,448 observations (the number of rows for each column). Also, please note that the `table() function` shows that every survey participant participated in the full survey. 

```{r}
# Import 

data <- read.dta(here::here("raw_data","naas16post-weighted-recoded.dta"))

# Participation 
table(data$fullsurvey)
```

# Turn factors into numbers

```{r}
# Turn factors into numbers 
data <- data %>%
  mutate(across(where(is.factor), compose(parse_number, as.character)))

# There are 569 variables 
ncol(data) 

names(data)
```

# Select and rename variables 

```{r}
# Create each dataframe
co_vars <- data %>%
    dplyr::select(
  
        # ID
        respid,     
       
        # Covariates
        male, # male 
        age, # age 
        educout_degree, # edu 
        ownhome, # home ownership 
        race, # race 
        democrat, # democrat
        republican, # republican
        q10_15, # income level
        statesmp, # states 
        forborn, # foreign born 
        citizen) # citizen

co_vars <- co_vars %>%
  rename("income" = "q10_15",
         "states" = "statesmp")

mar_vars <- data %>% dplyr::select(matches("micro|discrim"))

id_vars <- data %>% dplyr::select(matches("id_|id2_"))

linked_vars <- data %>% dplyr::select(matches("linked"))

comm_vars <- data %>% dplyr::select(matches("comm"))

# Bind them together
df <- reduce(list(co_vars, mar_vars, id_vars, linked_vars, comm_vars), bind_cols)

# Inspect
glimpse(df)
```

# Recode values 

## Some variables better treated as factors 

Some variables are better treated as factors because they are categorical variables (e.g., race). 

```{r}
# Race
df$race <- df$race %>%
    as.character() %>%
    recode("1" = "AAPI",  # Originally Asian Americans 
           "2" = "AAPI",  # Originally Pacific Islanders
           "3" = "White",
           "4" = "Black",
           "5" = "Black", # Originally African American
           "6" = "Latino",
           "9" = "Multiracial") %>%
    as.factor()

df <- df %>%
  mutate(race = as.factor(race),
         democrat = as.factor(democrat),
         republican = as.factor(republican),
         states = as.factor(states),
         forborn = as.factor(forborn),
         citizen = as.factor(citizen))
```
## Replace non-response values (`9`, `88`, `99`, `8888`, `9999`) with `NA`s

```{r}
df[df %in% c(9, 88, 99, 8888, 9999)] <- NA

glue("What's the percentage of missing values in the data AFTER replacing non-reponse values with NAs?: {(mean(is.na(df)) %>% round(2)) * 100}%")
```

## Rescale responses 

Rescale ordinal responses on the same scale (0-1).

```{r}
# Re-scale ordinal responses
rescaled <- df %>%
  dplyr::select(where(is.numeric)) %>% # starts with q
  apply(2, scales::rescale) %>%
  as.data.frame() 
```

```{r}
combined <- bind_cols(rescaled, df %>% select(-where(is.numeric)))
```

# Save the cleaned file 

```{r}
write.csv(combined, here("processed_data/cleaned.csv"))
```