---
title: "Imputation"
author: "Jae Yeon Kim"
output:
  html_document: 
    toc: true
    theme: united
---

# Setup 

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
        remotes, # for installing packages on GitHub 
        mice, # for imputation
        miceadds, # additional functions for mice
        tidyverse, # for the tidyverse framework 
        naniar, # for missing values 
        ggthemes, # for fancy ggplot themes
        here, # for self-contained projects
        ggpubr # for pub-ready themes 
)

source(here("functions", "utils.r"))

remotes::install_github("njtierney/naniar")

# for publication-friendly theme 
theme_set(theme_pubr())
```

# Importing files 

Importing the file that we created in `01_data_cleaning.Rmd`.

```{r}
# Import 
cleaned <- read.csv(here("processed_data/cleaned.csv"))

cleaned$X <- NULL
```

# Examining missing responses

- Listwise deletion (`na.rm = TRUE`) works only if these observations are missing completely at random (MCAR). This is a strong assumption, as it happens very rarely. 

## Exploratory analysis (data visualization)

- The `nanair package` provides many valuable functions for inspecting missing data values. I used the `miss_var_summary() function` from this package to check the missing pattern and visualized it using ggplot2. 

```{r}
# Visualize the percentage of missing values by variables and subgroups 

cleaned %>%
    group_by(race) %>% # group by race
    miss_var_summary() %>% # summary 
    filter(n_miss != 0) %>% # show variables that include missing values
    ggplot(aes(x = reorder(variable, pct_miss), y = pct_miss)) +
        geom_col() +
        facet_wrap(~race) + # facet by race 
        coord_flip() + # flip axes 
        labs(x = "Variables",
             y = "Missing rate (%)")
ggsave(here("outputs/missing_rate.png"), height = 8)
```

# Multiple imputation 

In this case, I took a multiple imputation approach. I recovered (imputed) missing values based on the observed data. I did so by making simulations of these imputed values to garner some uncertainty measures.

## Imputing data 

In this project, I used the `mice package` developed by [Stef van Buuren](https://stefvanbuuren.name/).  The following code shows how the imputation model is set up. I comment on each argument to make it explicit how the model is set up. For instance, the `m argument` refers to the number of imputations, and I set it to `5`.

```{r}
# Columns to exclude from imputation:
# - challeng_* has ~50% structural missingness by design (rotating block)
# - ethnic indicators are complete binary 0/1 with no missingness
no_impute_cols <- c(
  grep("^challeng_", names(cleaned), value = TRUE),
  "chinese", "filipino", "indian", "japanese", "korean",
  "vietnamese", "cambodian", "hmong", "bangladeshi",
  "pakistani", "hawaiian"
)

# Separate held-out columns from imputable columns
held_out <- cleaned %>% select(all_of(no_impute_cols))
to_impute <- cleaned %>% select(-respid, -all_of(no_impute_cols))

imp <- mice(
  to_impute,
    seed = 1234, # for reproducibility
    m = 5, # the number of imputations
    maxit = 100, # the max number of iterations
    method = "pmm", # predictive mean method
    print = FALSE)
```

## Pooling results

```{r}
# Pooling results
imputed <- mice::complete(imp, "long", include = FALSE) # save all 5 imputations

# Check
paste("The number of missing values in the imputed subset is", sum(is.na(imputed)))
```
 
# Reassemble and save the imputed data
 
```{r}
# We need to merge held_out columns back to the long format imputed data
# The imputed data has .imp and .id columns. We need to match by .id (which corresponds to row number in original)

# Add row number to held_out to facilitate join
held_out$id_row <- 1:nrow(held_out)

imputed <- imputed %>%
  rename(id_row = .id) %>%
  left_join(held_out, by = "id_row") %>%
  select(-id_row) # remove temporary id

paste("challeng_ vars still have NAs (by design):",
      sum(is.na(imputed %>% select(starts_with("challeng_")))))

write.csv(imputed, here("processed_data/imputed.csv"), row.names = FALSE)

names(imputed)
```
